An investigation into using scagnostics to find interesting projections of high-dimensional data

### 2018-10-23 scagnostics.Rmd
Make sure scales for rotation traces include 0, 1, finer binning, saved output.

## Rotation dependence
\label{sec:rot}
\textcolor{red}{testin rotation dependence by turning sine dataset with $\alpha \in [0,2\pi]$}
Interesting to note that some indexes are symmetric around $\pi$, but scagnostics indexes are very arbitrary, dependence is introduced via binnibng (and outlier removal?) here. Observe huge drops for some of the indexes! 
```{r rotationDep, fig.width=6, fig.height=7, fig.cap="Rotation dependence of the considered index functions tested on the sine 1000 dataset, by rotating the special view by an angle alpha."}
sineM <- as.matrix(select(sin1000, V5, V6))
sc <- tibble(
    holes=numeric(),
    cmass=numeric(),
    convex=numeric(),
    skinny=numeric(),
    stringy=numeric(),
    dcor2d=numeric(),
    splines2d=numeric(),
    MIC=numeric(),
    TIC=numeric(),
    alpha=numeric()
)
for (a in seq(0,2*pi, pi/20)){
  rotM <- matrix(c(cos(a), sin(a), -sin(a), cos(a)), ncol = 2)
  dprj <- sineM %*% rotM
  scagRes <- scagnostics(dprj)
  dcorRes <- dcor2d(dprj[,1], dprj[,2])
  splineRes <- splines2d(dprj[,1], dprj[,2])
  mineRes <- mine(dprj[,1], dprj[,2])
  holesRes <- holes(dprj)
  cmassRes <- cmass(dprj)
  sc <- add_row(sc, holes=holesRes, cmass=cmassRes,
                convex=scagRes[,"Convex"], skinny=scagRes[,"Skinny"],
                stringy=scagRes[,"Stringy"], dcor2d=dcorRes,
                splines2d=splineRes, MIC=mineRes$MIC,
                TIC=mineRes$TIC, alpha=a)
  }

scMelt <- melt(sc, id=c("alpha"))
ggplot(scMelt, aes(x=alpha, y=value)) +
  geom_line() + 
  facet_wrap(~variable, scales = "free_y", ncol = 1)
```

>>>>

## Rotation dependence
\label{sec:rot}
\textcolor{red}{testing rotation dependence by turning sine dataset with $\alpha \in [0,2\pi]$}
Interesting to note that some indexes are symmetric around $\pi$, but scagnostics indexes are very arbitrary, dependence is introduced via binnibng (and outlier removal?) here. Observe huge drops for some of the indexes! 
```{r rotationDep, fig.width=6, fig.height=7, fig.cap="Rotation dependence of the considered index functions tested on the sine 1000 dataset, by rotating the special view by an angle alpha."}
if(!file.exists("cache/rotationDep.rda")){
  sineM <- as.matrix(select(sin1000, V5, V6))
  sc <- tibble(
      holes=numeric(),
      cmass=numeric(),
      convex=numeric(),
      skinny=numeric(),
      stringy=numeric(),
      dcor2d=numeric(),
      splines2d=numeric(),
      MIC=numeric(),
      TIC=numeric(),
      alpha=numeric()
  )
  for (a in seq(0,2*pi, pi/100)){
    rotM <- matrix(c(cos(a), sin(a), -sin(a), cos(a)), ncol = 2)
    dprj <- sineM %*% rotM
    scagRes <- scagnostics(dprj)
    dcorRes <- dcor2d(dprj[,1], dprj[,2])
    splineRes <- splines2d(dprj[,1], dprj[,2])
    mineRes <- mine(dprj[,1], dprj[,2])
    holesRes <- holes(dprj)
    cmassRes <- cmass(dprj)
    sc <- add_row(sc, holes=holesRes, cmass=cmassRes,
                  convex=scagRes[,"Convex"], skinny=scagRes[,"Skinny"],
                  stringy=scagRes[,"Stringy"], dcor2d=dcorRes,
                  splines2d=splineRes, MIC=mineRes$MIC,
                  TIC=mineRes$TIC, alpha=a)
  }
  save(sc, file = "cache/rotationDep.rda")
} else {
  load("cache/rotationDep.rda")
}

scMelt <- melt(sc, id=c("alpha"))
ggplot(scMelt, aes(x=alpha, y=value)) +
  geom_line() + 
  facet_wrap(~variable, scales = "free_y", ncol = 1) +
  geom_blank(aes(y = 0)) +
  geom_blank(aes(y = 1))
```
